---
title: "Daf_proj"
author: '*'
date: "9/15/2021"
output: html_document
---
#EFFECT OF DISASTERS ON THE INSURANCE STOCK 

#Short Description:- 
In this project we are considering four datasets, the first dataset is "Disasters", in this dataset we are considering 5 disasters - Fire,Flood,Hurricane,Severe Stroms and Snow. The second,third and fourth datasets are Stocks data for top 3 listed insurance companies - AllState,Progressive and Travellers. So taking stocks data into consideration we are analyzying how disasters effected the stocks prices from year 2000 -2020

#Source:-
FEMA(Federal Emergency Managemnet Agency) Disaster Declarations Summary is a summarized dataset describing all federally declared disasters. This dataset lists all official FEMA Disaster Declarations, beginning with the first disaster declaration in 1953 and features all three disaster declaration types: major disaster, emergency, and fire management assistance. 
'https://www.fema.gov/api/open/v2/DisasterDeclarationsSummaries.csv'
Yahoo Finance - This is a media property that is part of Yahoo! netwwok.It provides financial news, data and commentary including stock quotes, press releases, financial reports, and original content.
'https://query1.finance.yahoo.com/v7/finance/download/ALL?period1=944006400&period2=1631664000&interval=1d&events=history&includeAdjustedClose=true'
'https://query1.finance.yahoo.com/v7/finance/download/PGR?period1=944006400&period2=1631664000&interval=1d&events=history&includeAdjustedClose=true'
'https://query1.finance.yahoo.com/v7/finance/download/TRV?period1=944006400&period2=1631664000&interval=1d&events=history&includeAdjustedClose=true'

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r importing tidyverse library}
library(tidyverse)
```

#Loading Disasters dataset
```{r disaster dataset}
disaster_data <- read_csv('https://www.fema.gov/api/open/v2/DisasterDeclarationsSummaries.csv')
```

#Loading AllState Stock dataset
```{r AllState dataset}
allstate_data <- read_csv('https://query1.finance.yahoo.com/v7/finance/download/ALL?period1=944006400&period2=1631664000&interval=1d&events=history&includeAdjustedClose=true')
allstate_data <- subset(allstate_data, select = c("Date","Open","High","Low","Close"))
colnames(allstate_data) <- c("Date", "allstate_open", "allstate_high","allstate_low","allstate_close")
head(allstate_data)
```
Selected the columns that are needed for the visualizations and formed a subset using those columns. Renamed the columns according to convenience


#Loading Progressive Stock dataset
```{r Progressive dataset}
progressive_data <- read_csv('https://query1.finance.yahoo.com/v7/finance/download/PGR?period1=944006400&period2=1631664000&interval=1d&events=history&includeAdjustedClose=true')
progressive_data <- subset(progressive_data, select = c("Date","Open","High","Low","Close"))
colnames(progressive_data) <- c("Date", "progressive_open", "progressive_high","progressive_low","progressive_close")
head(progressive_data)
```
Selected the columns that are needed for the visualizations and formed a subset using those columns. Renamed the columns according to convenience


#Loading Travellers Stock dataset
```{r Travellers dataset}
travellers_data <- read_csv('https://query1.finance.yahoo.com/v7/finance/download/TRV?period1=944006400&period2=1631664000&interval=1d&events=history&includeAdjustedClose=true')
travellers_data <- subset(travellers_data, select = c("Date","Open","High","Low","Close"))
colnames(travellers_data) <- c("Date", "travellers_open", "travellers_high","travellers_low","travellers_close")
head(travellers_data)
```
Selected the columns that are needed for the visualizations and formed a subset using those columns. Renamed the columns according to convenience


#Loading Disaster dataset
```{r}
disaster_data <- subset(disaster_data, select = c("incidentType", "state", "declarationType","fyDeclared", "declarationDate", "incidentBeginDate", "incidentEndDate", "disasterCloseoutDate"))
disaster_data <- disaster_data %>%  distinct()
disaster_data$incidentBeginDate = as.Date(disaster_data$incidentBeginDate)
disaster_data <- disaster_data %>% filter(incidentBeginDate>as.Date("2000-01-01"))
```
Selected the columns that are needed for the visualizations and formed a subset using those columns. To remove duplicate values distinct() function is used and filtered the values from the year 2000


#Combinning Stock and Disaster datasets
```{r join_data}
disaster_stock_data <- disaster_data %>% 
  left_join(allstate_data, by=c("incidentBeginDate" = "Date")) %>%
  left_join(progressive_data, by=c("incidentBeginDate" = "Date"))%>%
  left_join(travellers_data, by=c("incidentBeginDate" = "Date"))%>%
  na.omit((disaster_data_all))
disaster_stock_data
```
Joining all the four datasets to get into a single dataset, join is based on the date

```{r}
library(janitor)
summ <- tabyl(disaster_data,incidentType)
summ <- summ %>% filter(n >50)
summ %>% arrange(desc(n))
```
Till now we have loaded the datasets and filtered the data according to us ,
The Number of Disasters occurred on each year
```{r}
count_incidents <-tabyl(disaster_data, incidentType, fyDeclared)
count_incidents <- disaster_data %>% group_by(fyDeclared)%>%  summarise(number = n())
count_incidents
```
```{r}
ggplot(data=count_incidents, aes(x=fyDeclared, y = number))+ ggtitle("Total disasters occured in each year")+
    geom_line(color = "blue", linetype = 2)
```


What is the distribution of the natural calamities in US from 2000 - Present
```{r piechart}
library(ggrepel)
ggplot(summ %>% arrange(desc(n)), aes(x="", y=n, fill=incidentType ,inherit.aes = FALSE)) +
 geom_bar(stat="identity", width=1, color="white") +
 coord_polar("y", start=0) +
 theme_void() + ggtitle("Distribution of natural calamities: 2000 - now")
```

#Impact of natural calamities on the three Insurance Company Stocks
```{r}
disaster_stock_data<- disaster_stock_data %>% mutate(dip_allstate = allstate_close - allstate_open)
disaster_stock_data<- disaster_stock_data %>% mutate(dip_progressive = progressive_close- progressive_open )
disaster_stock_data<-disaster_stock_data %>% mutate(dip_travellers = travellers_close - travellers_open)
disaster_stock_data <- disaster_stock_data %>% mutate(dip = apply(disaster_stock_data[,c('dip_allstate','dip_progressive','dip_travellers')], 1, min, na.rm = TRUE))
glimpse(disaster_stock_data)

```

```{r}
list <- disaster_stock_data %>% filter(incidentType == c('Fire','Flood','Hurricane','Severe Storm(s)','Snow'))%>% group_by(incidentType) %>% summarise(mean_as = mean(dip_allstate), mean_tr = mean(dip_travellers), mean_pro = mean(dip_progressive))
summ <- full_join(summ,list)
summ
```
```{r}
summ_p <- pivot_longer(summ,cols = c('mean_as','mean_tr','mean_pro'))
ggplot(summ_p, aes(fill=name, y=value, x=incidentType)) + 
    geom_bar(position="stack", stat="identity") + ggtitle("Insurance Stock reactions to Disasters") + theme(legend.position="right",
          axis.text.x = element_text(angle = 90))

```
#Basic Linear Model of the data
```{r}
model <- lm(formula = n ~ mean_as+mean_tr+ mean_pro, data = summ)
summary(model)
```
#Conclusions:-
The most occurred disaster is Fire.
Highest number of disasters occcurred in 2011 whereas least number of disasters occurred in 2021
During Fire, Hurricane and Snow we have seen a raise in the Stocks for All State and Progressive Insurance companies.
In the time of flood and severe storms all the stocks of the three Insurance Companies are dropped.

#Bias:-
We considered the disasters that occurred more than 50 times.
Weekends are not considered as the stock markets are closed on weekends.





